---
title: "umiAnalyzer vignette"
author: "Stefan Filges"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{umiAnalyzer vignette}
  %\usepackage[utf8]{inputenc}
---

```{r, echo = FALSE}
  knitr::opts_chunk$set(collapse = TRUE, 
                        comment = "#>",
                        fig.width=9, 
                        fig.height=6)
```

# Introduction

umiAnalyzer provides tools for analyzing umierrorcorrect output. 

## How to make your own umiExperiment object

Define a variable containing the path to the directory with all the umierrorcorrect output folders 
belonging to your experiment. umiAnalyzer comes with raw test data generated with umierrorcorrect that 
you can import if you don't have any of your own.

Call the create.UMIexperiment to create your umiExperiment object.

The UMIexperiment object always maintains your raw data, however you can create as many filters as you
like, which will be saved as separate objects to access. You can filter the consensus table of
UMIexperiment object with filterUMIobject. The only mandatory arguments are the object to be filtered
and a user defined name. You can use that name to retrieve a filtered table using getFilter. 

```{r example1, eval=FALSE}
library(umiAnalyzer)

main <- system.file("extdata", package = "umiAnalyzer")

sample.names <- list.dirs(path = main, full.names = FALSE, recursive = FALSE)

simsen <- createUMIexperiment(
  experiment.name = "test",
  main.dir = main,
  dir.names = sample.names,
  importBam = FALSE
)

# save(simsen, file = "data/simsen.RData")

reads <- parseBamFiles(main.dir = main)

simsen <- generateQCplots(simsen, do.plot = TRUE, group.by = "assay")

simsen <- filterUMIobject(
  object = simsen, name = "myfilter", minDepth = 3,
  minCoverage = 100, minFreq = 0, minCount = 0
)

myfilter <- getFilter(object = simsen, name = "myfilter")
myfilter

simsen <- generateAmpliconPlots(object = simsen, filter.name = "myfilter", do.plot = TRUE)

metaData <- system.file("extdata", "metadata.txt", package = "umiAnalyzer")
simsen <- importDesign(object = exp1, file = metaData)

simsen <- merge_technical_replicates(object = simsen, filter.name = "myfilter")
simsen@merged.data

viz_Merged_data(simsen)
```

## Calling variants

umiAnalyzer comes with a build-in umiExperiment object to explore, which was generated using the code 
above, so it can be used without creating the it first if so desired.

In order to call variants using the umiAnalyzer variant caller simply load the package and test data
and use the callVariants function. You can then filter the resulting consensus data (cons.data) within
the object, e.g. for significant variants.

```{r example2, eval=FALSE}
#data <- simsen
#data <- callVariants(data)

#data <- filterVariants(object = data, p.adjust = 0.2, minDepth = 5)
```

## Importing experimental designs and statistics
### Experimental design
umiAnalyzer supports adding meta data to a UMIsample or UMIexperiment object, such as experimental 
design matrices or clinical parameters. This is done using the importDesign function and requires a 
simple formatted table supplied by the user as a tab separated file. It is important that the order 
of the samples in the meta data file is the same as when building the UMIexperiment object.

```{r design, eval=FALSE}
data <- simsen
data <- callVariants(data)

metaData <- system.file("extdata", "metadata.txt", package = "umiAnalyzer")

data <- importDesign(object = data, file = metaData)
```

### Working with meta data

It is also possible to add meta data to an object and to retrieve metadata if needed. The design 
matrix loaded with importDesign can be retrieved as follows:

```{r getmetadata, eval=FALSE}
design <- getMetaData(object = data, attributeName = "design")

design
```

Similarly, any kind of meta data can be added and retrieved from an object using addMetaData:

```{r addmetadata, eval=FALSE}
comment <- "fix this"
data <- addMetaData(object = data, attributeName = "my-comment", attributeValue = comment)

myattribute <- getMetaData(object = data, attributeName = "my-comment")
myattribute
```




