---
title: "umiAnalyzer vignette"
author: "Stefan Filges"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{umiAnalyzer vignette}
  %\usepackage[utf8]{inputenc}
---

```{r, echo = FALSE}
  knitr::opts_chunk$set(collapse = TRUE, 
                        comment = "#>")
```

# Introduction

umiAnalyzer provides tools for analyzing umierrorcorrect output. 

## How to make your own umiExperiment object

Define a variable containing the path to the directory with all the umierrorcorrect output folders 
belonging to your experiment. umiAnalyzer comes with raw test data generated with umierrorcorrect that 
you can import if you don't have any of your own.

Call the create.UMIexperiment to create your umiExperiment object.

```{r example1}
library(umiAnalyzer)

main = system.file("extdata", package = "umiAnalyzer")

sample.names <- list.dirs(path = main, full.names = FALSE, recursive = FALSE)

exp1 <- create.UMIexperiment(experiment.name = "exp1",
                             main.dir = main,
                             dir.names = sample.names)
```

## A basic example

umiAnalyzer comes with a build-in umiExperiment object to explore, which was generated using the code 
above, so it can be used without creating the it first if so desired.

In order to call variants using the umiAnalyzer variant caller simply load the package and test data
and use the callVariants function. You can then filter the resulting consensus data (cons.data) within
the object, e.g. for significant variants.

```{r example2}
data <- simsen
data <- callVariants(data)

vars <- data@cons.data
vars <- vars[vars$p.adjust <= 0.1,]
```

## Importing experimental designs and statistics
### Experimental design
umiAnalyzer supports adding meta data to a UMIsample or UMIexperiment object, such as experimental design matrices or clinical parameters. This is done using the importDesign function and requires a simple formatted table supplied by the user
as a tab separated file. It is important that the order of the samples in the
meta data file is the same as when building the UMIexperiment object.

```{r design}
data <- simsen
data <- callVariants(data)

metaData <- system.file("extdata", "metadata.txt", package = "umiAnalyzer")

data <- importDesign(object = data, file = metaData)

attributes(data)$design

```

### General addition of meta data





